---
title: "California Oil Spills Spatial Data Analysis"
description: |
  My Spatial Visualization Skills on Display.
author:
  - name: Kiera Matiska
date: 2022-03-17
output:
  distill::distill_article:
    self_contained: false
---

# Overview

This code will provide information on oil spills and their location throughout California State. A t-map will be produced as an interactive mode of exploring the data and a second static choropleth graph will be used to show a larger picture of the data within each county.

### Citation

California Department of Fish and Game & Office of Spill Prevention and Response. (20089, July 23). *Oil Spill Incident Tracking.* Digital map and vector digital data. https://map.dfg.ca.gov/metadata/ds0394.html#ID0EUGA 

## Set-Up

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach packages
library(tidyverse)
library(here)
library(maptools)
library(sf)
library(tmap)
library(tmaptools)
library(janitor)
```

## California Oil Spill Maps

### T-map

```{r}
oil_spills_sf <- read_sf(here("data/ds394/ds394.shp")) %>% # read in shape file
  clean_names() %>% 
  st_transform(3310) # transform coordinate system

ca_counties <- read_sf(here("data/CA_Counties/CA_Counties_TIGER2016.shp")) %>% # read in shape file
  clean_names() %>% 
  st_transform(3310) # transform coordinate system

tmap_mode(mode = "view") # interactive t-map option
tmap_options(check.and.fix = TRUE,
             max.categories = 58)

tm_shape(ca_counties) + # start by using CA counties
  tm_fill("aland", # fill by land area
          palette = "BuGn", 
          popup.vars = c("County:" = "namelsad")) + # add a popup variable
  tm_borders(col = "black") + # add borders around the counties
tm_shape(oil_spills_sf) + # add another layer
  tm_dots(popup.vars = c("Control Number:" = "oesnumber", # add more popup variables
                         "Date:" = "dateofinci", 
                         "Time:" = "timeofinci", 
                         "County:" = "localecoun", 
                         "City:" = "localecity", 
                         "Type of Water:" = "specificlo",
                         "Inland or Marine:" = "inlandmari"),
          size = 0.02) # change size of points
```

**Figure 1.** Interactive Map Showing Locations of Oil Spills within California State. For each oil spill location, the user may click on the point to receive more information including the data, time, and location.

### Choropleth Map

```{r}
ca_oil <- ca_counties %>% 
  st_join(oil_spills_sf) %>% # join the two data sets together
  select(latitude, longitude, namelsad) %>% 
  group_by(namelsad) %>% # group by name of county
  summarize(oil_spills = n()) # count the number of oil spills in each county

ggplot(data = ca_oil) + 
  geom_sf(aes(fill = oil_spills)) + # define sf plot and color by oil_spill count
  scale_fill_gradient(low = "steelblue1", high = "dodgerblue4") + # change colors
  labs(title = "Choropleth Map of Oil Spills by \nCounty in California State", # add title
       x = "Longitude", # add x-axis label
       y = "Latitude", # add y-axis label
       fill = "Oil Spill Number") + # change gradient legend name
  theme(plot.title = element_text(hjust = 0.5)) # center title
```

**Figure 2.** Choropleth Map of Oil Spills by County in California State. The colors indicate the number of oil spills that have occurred in each county.

