---
title: "Text Analysis of the Musical: Hamilton"
description: |
  My Text Analysis Skills on Display.
author:
  - name: Kiera Matiska
date: 2022-03-15
output:
  distill::distill_article:
    self_contained: false
---

# Introduction

On July 13, 2015, the musical *Hamilton*, written by Lin-Manuel Miranda, premiered at the Richard Rogers Theater on Broadway. It received critical acclaim and the musical was nominated 16 times for Tony award categories and won in 11 categories. The musical was also awarded the 2016 Pulitzer Prize for Drama. It instantly became a smashing hit, with people all around the globe not only listening to the music, but striving to see it in person. 

The musical tells the story of the Founding Father Alexander Hamilton, with the music heavy on hip-hop, R&B, rap, and soul. It is split into two acts that explore his life before becoming a major political figure and after. The musical encounters other important historical characters, such as George Washington, Thomas Jefferson, and Aaron Burr. 

Here is the original broadway cast:

- Alexander Hamilton: Lin-Manuel Miranda
- Aaron Burr: Leslie Odom Jr.
- Eliza Hamilton: Phillipa Soo
- Angelica Schuyler: Renée Elise Goldsberry
- Marquis de Lafayette & Thomas Jefferson: Daveed Diggs
- George Washington: Christopher Jackson
- John Laurens & Philip Hamilton: Anthony Ramos
- Hercules Mulligan & James Madison: Okieriete Onaodowan
- King George III: Johnathon Groff
- Peggy Schuyler & Maria Reynolds: Jasmine Cephas Jones

This code will provide a text analysis on the lyrics of each song between the two acts. Word counts will be assessed, the top 5 most frequently used words per song will be graphed. Additionally, two word clouds will be generated based on the most frequent 100 words in each act. Finally, a sentiment analysis will be ran, comparing the attitudes and opinions expressed throughout each song between the two acts.

## Set Up

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# attach packages
library(tidyverse)
library(here)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
library(paletteer)
```

## Wrangle Data

```{r, cache = TRUE}
## read in lyrics for act 1 and 2
ham_act1 <- pdf_text(here("data", "hamilton_act1.pdf")) # read in act 1 lyrics

ham_act2 <- pdf_text(here("data", "hamilton_act2.pdf")) # read in act 2 lyrics

## convert text to a data frame
act1_lines <- data.frame(ham_act1) %>% # create a data frame using act 1 lyrics
  mutate(page = 1:n()) %>% # define all pages in the text
  mutate(text_act1 = str_split(ham_act1, pattern = "\\n")) %>% # create a new row 
                                                               # with every new line
  unnest(text_act1) %>% 
  mutate(text_act1 = str_trim(text_act1))

act2_lines <- data.frame(ham_act2) %>% # create a data frame using act 2 lyrics
  mutate(page = 1:n()) %>% # define all pages in the text
  mutate(text_act2 = str_split(ham_act2, pattern = "\\n")) %>% # create a new row 
                                                               # with every new line
  unnest(text_act2) %>% 
  mutate(text_act2 = str_trim(text_act2))

## do some tidying
tunes_act1 <- act1_lines %>% # create a tunes data frame
  mutate(song = ifelse(str_detect(text_act1, "Song"), text_act1, NA)) %>% # create a song column
  fill(song, .direction = "down") %>% # fill down to create rows
  separate(col = song, into = c("so", "no"), sep = " ") %>% # create new columns with song 
                                                            # in one and number in the other
  mutate(song = as.numeric(as.roman(no))) # number column as roman numeral

tunes_act2 <- act2_lines %>% # create a tunes data frame
  mutate(song = ifelse(str_detect(text_act2, "Song"), text_act2, NA)) %>% # create a song column
  fill(song, .direction = "down") %>% # fill down to create rows
  separate(col = song, into = c("so", "no"), sep = " ") %>% # create new columns with song 
                                                            # in one and number in the other
  mutate(song = as.numeric(as.roman(no))) # number column as roman numeral
```

## Find Word Counts and Remove Stop Words

```{r}
# unnest the words in each column of the act 1 text
words_act1 <- tunes_act1 %>% 
  unnest_tokens(word, text_act1) %>% 
  select(-ham_act1)

# unnest the words in each column of the act 2 text
words_act2 <- tunes_act2 %>% 
  unnest_tokens(word, text_act2) %>% 
  select(-ham_act2)

# find word count of act 1 separated by song
act1_wordcount <- words_act1 %>% 
  count(song, word)

# find word count of act 2 separated by song
act2_wordcount <- words_act2 %>% 
  count(song, word)

# Remove stop words in act 1
words_act1_clean <- words_act1 %>% 
  anti_join(stop_words, by = "word") %>% 
  mutate(song_name = case_when( # rename songs in act 1
    song == "1" ~ "Alexander Hamilton",
    song == "2" ~ "Aaron Burr, Sir",
    song == "3" ~ "My Shot",
    song == "4" ~ "The Story of Tonight",
    song == "5" ~ "The Schuyler Sisters",
    song == "6" ~ "Farmer Refuted",
    song == "7" ~ "You'll Be Back",
    song == "8" ~ "Right Hand Man",
    song == "9" ~ "A Winter's Ball",
    song == "10" ~ "Helpless",
    song == "11" ~ "Satisfied",
    song == "12" ~ "The Story of \nTonight (Reprise)",
    song == "13" ~ "Wait for It",
    song == "14" ~ "Stay Alive",
    song == "15" ~ "Ten Duel Commandments",
    song == "16" ~ "Meet Me Inside",
    song == "17" ~ "That Would Be Enough",
    song == "18" ~ "Guns and Ships",
    song == "19" ~ "History Has Its Eyes on You",
    song == "20" ~ "Yorktown (The World \nTurned Upside Down)",
    song == "21" ~ "What Comes Next?",
    song == "22" ~ "Dear Theodosia",
    song == "23" ~ "Non-Stop"
  ))

# remove stop words in act 2
words_act2_clean <- words_act2 %>% 
  anti_join(stop_words, by = "word") %>% 
  mutate(song_name = case_when( # rename songs in act 2
    song == "1" ~ "What'd I Miss",
    song == "2" ~ "Cabinet Battle #1",
    song == "3" ~ "Take a Break",
    song == "4" ~ "Say No to This",
    song == "5" ~ "The Room Where It Happens",
    song == "6" ~ "Schuyler Defeated",
    song == "7" ~ "Cabinet Battle #2",
    song == "8" ~ "Washington on Your Side",
    song == "9" ~ "One Last Time",
    song == "10" ~ "I Know Him",
    song == "11" ~ "The Adams Administration",
    song == "12" ~ "We Know", 
    song == "13" ~ "Hurricane", 
    song == "14" ~ "The Reynolds Pamphlet", 
    song == "15" ~ "Burn",
    song == "16" ~ "Blow Us All Away", 
    song == "17" ~ "Stay Alive - Reprise", 
    song == "18" ~ "It's Quiet Uptown", 
    song == "19" ~ "The Election of 1800", 
    song == "20" ~ "Your Obedient Servant", 
    song == "21" ~ "Best of Wives and \nBest of Women", 
    song == "22" ~ "The World Was Wide Enough",
    song == "23" ~ "Who Lives, Who Dies, \nWho Tells Your Story"
  ))

# find word counts without stop words in act 1
act1_nonstop_counts <- words_act1_clean %>% 
  count(song, song_name, word)

# find word counts without stop words in act 2
act2_nonstop_counts <- words_act2_clean %>% 
  count(song, song_name, word)
```

## Find top 5 words from each song in each Act {.tabset .tabset-fade}

```{r}
# top 5 words in act 1
act1_top5_words <- act1_nonstop_counts %>%
  group_by(song) %>%
  arrange(-n) %>% # arrange by decreasing number
  slice(1:5) %>% # choose top 5 words in each song
  ungroup() # stop grouping by song

# top 5 words in act 2
act2_top5_words <- act2_nonstop_counts %>%
  group_by(song) %>%
  arrange(-n) %>% # arrange by decreasing number
  slice(1:5) %>% # choose top 5 words in each song
  ungroup() # stop grouping by song
```

### Act 1

#### Songs 1-6

```{r}
# graph the top 5 words in song 1-6
act1_1_6 <- act1_top5_words %>% 
  filter(song %in% c("1", "2", "3", "4", "5", "6")) # filter for songs 1-6

ggplot(data = act1_1_6, # specify data
       aes(x = n, y = fct_reorder(word, n))) + # define parameters
  geom_col(fill = "#f6c049", color = "#a8a7a7") + # define column graph and change colors
  facet_wrap(~song_name, scales = "free") + #separate by song
  labs(title = "Top 5 Words in Songs 1-6 in Hamilton: Act 1", # add title
       x = "Count", # add x-axis label
       y = "Word") + # add y-axis label
  theme(plot.title = element_text(hjust = 0.5), # center title
        panel.background = element_rect(color = "#404040"), # outline panel
        plot.background = element_rect(fill = "snow")) # color plot background
```
