---
title: "Sea Level Changes of Mono Lake in Lee Vining, California"
description: |
  My Data Visualization Skills on Display
author:
  - name: Kiera Matiska
date: 2021-11-27
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach packages
library(tidyverse)
library(here)
library(janitor)
library(scales)
library(dplyr)
```

```{r}
mono_lake <- read_csv(here("data", "mono_lake_annual_levels1.csv"), skip = 5) %>% # read in data
  clean_names() # changed column headers to lower_snake_case
```

```{r}
ggplot(data = mono_lake, # call ggplot to make a funtion using the mono_lake data frame
       aes(x = year, y = lake_level_feet_above_sea_level)) + # define x and y axis
  geom_point(color = "steelblue4") + # specify scatterplot and change point color
  scale_x_continuous(breaks = pretty_breaks(n = 16)) + # axis break every 10 years
  labs(title = "Lake Level Changes in Mono Lake from 1850 to 2017", # title label
       x = "Year", # x-axis label
       y = "Lake Level (feet above sea level)") + # y-axis label
  theme_light() + # changed theme
  theme(plot.background = element_rect(fill = "lightskyblue2"), # changed color
        panel.background = element_rect(fill = "white"), # changed color
        panel.grid.major = element_line(color = "gray77"), # changed color
        plot.title = element_text(hjust = 0.5), # centered title
        legend.position = "none") + # removed legend
  geom_segment(aes(x = 1958, y = 6427, xend = 1921, yend = 6427, color = "red"), # defined constraints
               arrow = arrow(length = unit(0.8, "mm"))) + # added upper arrow
  geom_segment(aes(x = 1942, y = 6374, xend = 1979, yend = 6372, color = "red"), # defined constraints
               arrow = arrow(length = unit(0.8, "mm"))) + # added lower arrow
  geom_segment(aes(x = 1939, y = 6410, xend = 1941, yend = 6416, color = "red"), # defined constraints
               arrow = arrow(length = unit(0.8, "mm"))) + # added first key date arrow
  geom_segment(aes(x = 1981, y = 6385, xend = 1978.5, yend = 6376, color = "red"), # defined constraints
               arrow = arrow(length = unit(0.8, "mm"))) + # added second key date arrow
  geom_segment(aes(x = 1996, y = 6395, xend = 1994, yend = 6375.6, color = "red"), # defined constraints
               arrow = arrow(length = unit(0.8, "mm"))) + # added third key date arrow
  geom_hline(yintercept = 6377) + # added horizontal line
  geom_hline(yintercept = 6392) + # added horizontal line
  annotate("text", x = 1967, y = 6427, 
           label = "High: 6,427 ft", size = 2.5) + # added text on graph
  annotate("text", x = 1931, y = 6374, 
           label = "High: 6,372.30 ft", size = 2.5) + # added text on graph
  annotate("text", x = 1935, y = 6408, 
           label = "1941: Lake water \n diverted to Los Angeles", size = 2.5) + # added text on graph
  annotate("text", x = 1983, y = 6388, 
           label = "1978: Mono Lake \n Committee formed", size = 2.5) + # added text on graph
  annotate("text", x = 1996, y = 6398, 
           label = "1994: Decision \n 1631 Passed", size = 2.5) + # added text on graph
  annotate("text", x = 1890, y = 6378.5, 
           label = "Land Bridge Forms", size = 2.5) + # added text on graph
  annotate("text", x = 1890, y = 6393.5, 
           label = "Decision 1631: Lake Level Compromise of 6,392 above Sea Level", size = 2.5) # added text on graph
```

**Figure 1.** Lake Level Changes in Mono Lake from 1850 to 2017. Located northeast of Yosemite National Park in Lee Vining, California, Mono Lake is home to a plethora of different plant and animal species. Throughout the recorded years, 1918 and 1919 had the highest lake level of 4,627 feet above sea level and 1981 had the lowest (around 6,372,30 feet above sea level). When considering wildlife distribution in the area, it is imperative to look at the lake level. Most years have lake levels upwards of 6,377 feet above sea level. When the water level is, at maximum, this high, land bridges are formed that allow predators to attack the bird species that occupy and nest on the island in the center of the lake. In 1941, water from this lake was diverted to Los Angeles to provide a source of clean water for the population there. This decreased the volume of the lake and resulted in the low lake level that allowed the land bridge to form. Once the Mono Lake Committee was formed in 1978, lake levels have started to rise again, minimizing the risk of the land bridge forming and protecting bird species nesting on the island.
