---
title: "Snowshoe Hare Analysis"
description: |
  Statistical Analysis on Different Variables Collected on Snowshoe Hares
author:
  - name: Kiera Matiska
date: 2022-02-06
output:
  distill::distill_article:
    self_contained: false
---

# Introduction

In Alaska, Bonanza Creek is populated with snowshoe hares. Snowshoe hares are a keystone species in northern boreal forests and are therefore crucial to other species living within the same habitat (Kielland et al., 2017). They are important as prey for larger ground and air animals and can also do a lot of damage to vegetation if their population gets too large (National Parks Service, 2021). Due to lack of quantitative descriptions of hare population changes, this data was collected via capture-recapture strategies to count the population of hares in this creek. In this report, I use statistical tests and analyses to describe the changes in juvenile hare populations over 12 years (from 1999 - 2012).

# Data and Analyses

The data set used in this report has 3197 observations of hare individuals between three sites. These sites are Bonanza Riparian, Bonanza Mature Forest, and Bonanza Lowland Mature Forest. Each observation contains data on the age, sex, weight (g), hind foot length (mm), site, left and right ear tags, date, and trap ID. For this report, calculations using t-tests (significance level: ($\alpha$) = 0.05) and linear regression test are used to determined difference between means and the relationship between hind foot length and weight, respectively. Additionally, a Cohen's *d* test is used to determine relationship effect size between hind feet length and weight and a Pearson's *r* tests to determine the variance in the same relationship. Data is represented clearly in a summary table (with mean weight, median weight, standard deviation, minimum and maximum weights, and sample size) and over a multitude of different graphs. Statistical analyses are done using R coding language in RStudio.

```{r setup, include = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# Attach Packages
library(tidyverse)
library(here)
library(lubridate)
library(kableExtra)
library(effsize)
library(broom)
```

```{r}
# Read in data
# - Use `read_csv()` and the here package to read in bonanza_hares.csv

hares <- read_csv(here("data", "bonanza_hares1.csv")) # reads in data
```

# 1. Annual Juvinile Hare Trap Counts

```{r}
# Data Wrangling

# - Create a new column, year, using the `year()` function in lubridate
# - Create a subset that only contains juvenile hare data
# - Create a new data frame using the `summarize()` function that has two columns, year and number of hare traps
# - Use `kable_styling()` to finalize the table

hares_year <- hares %>% # stores new subset
  mutate(date = mdy(date)) %>% # changes date column to class date
  mutate(year = year(date)) # adds a new column with just year

hares_juvenile <- hares_year %>% # stores new subset
  filter(age == "j") # filters out observations in the age column that is not "j"

hares_traps <- hares_juvenile %>% # stores a new subset
  select(year, age) %>% # selects only columns age and year
  group_by(year) %>% # groups the subset by year
  summarize(number = n()) # creates a summary table with only year and number of traps for juvenile hares

hares_traps %>% # finalize this table
  kbl(caption = "Table 1. Number of Juvenile Hare Traps Each Year from 1999 - 2012") %>% # adds a caption
  kable_styling()
```

```{r}
ggplot(data = hares_traps, # identifies data set
       aes(x = year, y = number)) + # sets x and y axis
  geom_col(color = "gray", fill = "darkblue") + # makes a column graph and adds color to data
  scale_x_continuous(breaks = 1999:2012) + # add each year label under observations
  labs(x = "Year", # titles x-axis
       y = "Number of Traps") + # titles y-axis
  theme(plot.background = element_rect(fill = "lavender"), # changes the plot background color
        panel.background = element_rect(fill = "gray94"), # changes the panel background color
        panel.grid.major = element_line(color = "snow3")) # change color of lines on the panel
```

**Figure 1.** Annual Counts of Juvenile Hare Traps in Bonanza Creek. Sample size is `r sum(hares_traps$number[1:12])` traps.


The maximum and minimum amount of Juvenile Hare Trappings are `r hares_traps$number[1]` (in 1999) and `r hares_traps$number[10]` (in 2009) respectively. The median number of traps is `r median(hares_traps$number)` and the mean value is `r mean(hares_traps$number)`. These two values show that there has been a substantial decrease in traps since 1999. Hence, after 1999, the general trend of juvenile traps is negative, meaning that overall, the trap counts decrease from the first year onward. On the other hand, these numbers are highly influenced by effort, or the amount of time the researchers put into trapping and re-trapping these hares. By documenting the number of days in the field for a single year and then repeating that annually may lead to clearer results.

# 2. Juvenile Hare Weights

```{r}
# Data Wrangling

# - Using `mutate()`, rename the sex variables so they are spelled out: m = male, f = female, 
# and NA = not applicable
# - Using `mutate()`, rename the grid variables so they are spelled out: 
# bonrip = Bonanza Riparian, bonmat = Bonanza Mature, bonbs = Bonanza Lowland Mature
# - Create new data frame, hare_weights, using the `select()` 
# function so that it contains only hare weights, sex, and grid site.

hares_sex <- hares_juvenile %>% # stores new subset
  mutate(sex = case_when( # renames variables in the sex column
    sex == "m" ~ "male", 
    sex == "f" ~ "female",
    is.na(sex) == TRUE ~ "N/A" # changes NA filler variables to N/A for not applicable
  ))

hare_grids <- hares_sex %>% # stores new subset
  mutate(grid = case_when( # renames variables in grid column
    grid == "bonrip" ~ "Bonanza Riparian",
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonbs" ~ "Bonanza Lowland Spruce"
  ))

hare_weights <- hare_grids %>% # stores a new subset
  select(weight, sex, grid) # with only these columns
```

```{r}
hare_weights_size1 <- hare_weights %>% 
  select(sex, grid, weight) %>% 
  group_by(grid, sex) %>% 
  summarize(sample_size = n(),
            mean = mean(weight, na.rm = TRUE))

hare_weights_size2 <- hare_weights %>% 
  select(sex, grid) %>% 
  group_by(grid) %>% 
  summarize(sample_size = n())
```

```{r}
ggplot(data = hare_weights, # identifies data set
       aes(x = sex, y = weight)) + # sets x and y variables
  geom_point(size = 0.8, aes(color = sex)) + # scatter plot with colors based on sex size change of each point
  facet_grid(~ grid) + # separate data on graph by site
  labs(x = "Sex", # x-axis label
       y = "Weight (g)") + # y-axis label
  theme(legend.position = "none", # removes legend
        plot.background = element_rect(fill = "lavender"), # changes the plot background color
        panel.background = element_rect(fill = "gray94"), # changes the panel background color
        panel.grid.major = element_line(color = "snow3")) # change color of lines on the panel
```

**Figure 2.** Juvenile Hare Weights (g) by Sex in Each Bonanza Creek Site Zone. The first site ('Bonanza Lowland Spruce') is the site used for monitoring abundance and survival of snowshoe hares via live-trapping them. The second site ('Bonanza Mature') consists of mature white spruce, alder, paper birch, and black spruce trees. The last site ('Bonanza Riparian') is the flood zone area of the creek nearby the Tanana River and the Bonanza Mature Black Spruce. The sample sizes of each of the three sites respectively are: `r hare_weights_size2$sample_size[1]` hares,  `r hare_weights_size2$sample_size[2]` hares, and `r hare_weights_size2$sample_size[3]` hares.


On average, female juvenile hares are smaller than male hares. In each site, male hares are roughly 15 g larger in site 1, 142 g in site 2, and 105 g in site 3. Hares are the largest, on average, in site 1. Additionally, each site has different size populations of juvenile hares based on this trapping data. The smallest population is in the Lowland Bonanza Mature site (`r hare_weights_size2$sample_size[1]` hares) and the largest population is in the Bonanza Riparian site (`r hare_weights_size2$sample_size[3]` hares).

# 3. Juvenile Weight Comparisons (male and female snowshoe hares)

```{r}
# Wrangle Data
# Create a data subset, `weight_hare` with only the variables sex and weight

weight_hare <- hares_sex %>% # stores a new subset
  select(sex, weight) %>% # with only these columns
  filter(sex %in% c("male", "female")) # without NA data
```

```{r}
hare_summary <- weight_hare %>% # stores a summary table
  group_by(sex) %>% # group by sex
  summarize(min_weight_g = min(weight, na.rm = TRUE), # adds min weight for both sexes
            max_weight_g = max(weight, na.rm = TRUE), # adds max weight for both sexes 
            mean_weight = round(mean(weight, na.rm = TRUE), 2), # mean the weight, remove NA values 
            median_weight_g = median(weight, na.rm = TRUE), # gives the middle value of both sexes
            sd = round(sd(weight, na.rm = TRUE), 2), # standard deviate the weight, remove NA values
            sample_size = n()) # calculate sample size

hare_summary %>% # finalize summary table
  kbl(col.names = c("Sex", # changes column headers
                    "Minimum Weight (g)",
                    "Maximum Weight (g)",
                    "Mean Weight (g)",
                    "Median Weight (g)",
                    "Standard Deviation",
                    "Sample Size"),
    caption = "Table 1: Mean, Standard Deviation, and Sample Sizes of Weights by Sex of Juvenile Hares") %>% # add a caption
  kable_styling()
```

```{r}
# Create vectors of hare weights 
weight_males <- hares_sex %>% # Stores a vector
  filter(sex == "male") %>% # with only male observations
  drop_na(weight) %>% # dropping weights with NA stored
  pull(weight) # separates weight values into vector

weight_females <- hares_sex %>% # stores a vector
  filter(sex == "female") %>% # with only male observations
  drop_na(weight) %>% # dropping weights with NA stored
  pull(weight) # separates weight values into vector
```

```{r, include = FALSE}
plot(weight_males) # Plots male weights

plot(weight_females) # Plots female weights

hist(weight_males) # create a histogram for male weights

hist(weight_females) # create a histogram for female weights

qqplot(weight_males, weight_females) # create a qqplot of male and female weights
```

```{r}
#Perform a t-test and cohen's d test on the data
# Use `t.test()` to run the test and find significance of the data. Use `cohen.d()` to run a cohens d test.

weight_ttest <- t.test(weight_males, weight_females) # conduct t-test

weight_cohen <- cohen.d(weight_males, weight_females) # conduct cohen d's test -> Small effect size
```

Between male and female juvenile hares, on average, male hares are larger than female hares by 91 grams. The effect size was small based on the Cohen's *d* value of `r weight_cohen$estimate` and the unpaired t-test resulted in a p-value of `r weight_ttest$p.value`. This p-value means that there is a `r (weight_ttest$p.value)*100`% chance of finding sample means that are at least this different if drawn, by random chance, from populations with the same mean weight.

# 4. Relationship Between Juvenile Weight and Hind Foot Length

```{r}
# Data Wrangling
# Create a new subset that consists of only juvenile hares, weights, and hind foot lengths in mm

weight_hind <- hares_juvenile %>% # stores new subset
  select(weight, hindft) # with only these columns
```

```{r}
ggplot(data = weight_hind, # select this data
       aes(x = hindft, y = weight)) + # selects x and y variables
  geom_point(color = "springgreen4") + # scatter plot with green points
  labs(x = "Weight (g)", # x-axis label
       y = "Hind Foot Length (mm)") + # y-axis label
  theme(plot.caption = element_text(hjust = 0), # positions figure label to the bottom left of the graph
        plot.background = element_rect(fill = "lavender"), # changes the plot background color
        panel.background = element_rect(fill = "gray94"), # changes the panel background color
        panel.grid.major = element_line(color = "snow3")) # change color of lines on the panel
```

**Figure 3.** Weight versus Hind Foot Length of Juvenile Hares. Weight is in grams and hind foot length is in millimeters. Sample size is 378 individuals.

```{r}
# Statistical Testing

weight_hind_lm <- lm(weight ~ hindft, data = weight_hind) # run a linear regression test

weight_hind_tidy <- tidy(weight_hind_lm) # Shows lm stats

weight_hind_glance <- glance(weight_hind_lm)

weight_hind_pearsons <- cor.test(weight_hind$weight, weight_hind$hindft) # Pearson's R Test
```

For this section on the relationship between weight (g) and hind foot length (mm), it appears that there is a positive correlation between hind foot length and weight. For the linear regression model test, the hind foot coefficient is `r round(weight_hind_lm$coefficients[2], 2)`. This value means that with every 1 mm increase in hind foot length, the weight of the hare increases by `r round(weight_hind_lm$coefficients[2], 2)` grams on average. Additionally, the R^2 value is `r round(weight_hind_glance$r.squared, 3)`. The meaning of this value is that `r round(weight_hind_glance$r.squared, 3)*100`% of variance in hare weight is explained by hind foot length. The Pearson's *r* value of `r round(weight_hind_pearsons$estimate, 2)` means that there is a moderate correlation between hind foot length and hare weight. The assumptions of linear regression that may be of concern from this is homoscedesticity because the variances are unequal meaning that the variances of this relationship is actually heteroscedastic.

# Summary

To summarize this report, we found:

- that hare traps decreased from 1999 to 2012, showing a possible decrease in effort of the trapping and re-capturing program of snowshoe hares
- that there was a higher abundance of hares among the Bonanza Riparian Site
- that female juvenile hares were, on average, slightly smaller than male juvenile hares
- that there is a moderate, positive correlation between hind foot length (mm) and snowshoe hare weight (g)

# Citations

David Robinson, Alex Hayes and Simon Couch (2021). broom: Convert Statistical Objects into Tidy Tibbles. R
package version 0.7.9. https://CRAN.R-project.org/package=broom

Garrett Grolemund, Hadley Wickham (2011). Dates and Times Made Easy with lubridate. Journal of Statistical
Software, 40(3), 1-25. URL https://www.jstatsoft.org/v40/i03/.

Hao Zhu (2021). kableExtra: Construct Complex Table with 'kable' and Pipe Syntax. R package version 1.3.4.
https://CRAN.R-project.org/package=kableExtra

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

Kirill Müller (2020). here: A Simpler Way to Find Your Files. R package version 1.0.1.
https://CRAN.R-project.org/package=here

National Parks Service. 2021. NETN Species Spotlight - Snowshoe Hare. National Parks Service. https://www.nps.gov/articles/netn-species-spotlight-snowshoe-hare.htm

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical
Computing, Vienna, Austria. URL https://www.R-project.org/.

RStudio Team (2021). RStudio: Integrated Development Environment for R. RStudio, PBC, Boston, MA URL
http://www.rstudio.com/.

Torchiano M (2020). _effsize: Efficient Effect Size Computation_. doi: 10.5281/zenodo.1480624 (URL:
https://doi.org/10.5281/zenodo.1480624), R package version 0.8.1, <URL:
https://CRAN.R-project.org/package=effsize>.

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686,
https://doi.org/10.21105/joss.01686
